---
title: "Project Checkpoint 3"
author: "Phuc Nguyen, Jina Park"
date: "November 8, 2017"

output: html_document
---

# Progress Made

- Created a Github repository for our R project.

- Cleaned the Chicago crime data by filtering out dates to only include crimes committed between 2010-2017, changed the categorical variables of Arrest and Domestic to binary variables, and changed the Date column from strings to datetime format.

- Obtained a Census API key and demographic data using acs and function called ‘get_tract_demographics.’

- Started looking up how to create maps using choropleth and choroplethrMaps packages


# Roles of each member

- Phuc wrote up the meanings of the variables in the Chicago crime data.

- Phuc typed up the commands to clean the data (i.e. change Date column to datetime format, change categorical variables to dummy variables, etc).

- Jina wrote up the meanings of the variables in the demographic data.

- Jina worked with choroplethr and acs to obtain the demographic data.

- Jina typed up this report.

- We both worked to set up a Github repository so that we could share the R files.

- We both worked together during the entire process and helped each other through most of the steps.


# Chicago Crime Data

### Description

This data set contains information about crime incidents from 2010 to present (2017) in Chicago.

We obtained this data from [https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2/data]

Some variables that we might use in our analysis are

- ID: unique identifier for each record

- Date: the best estimate of when (date, time) the incident occurred

- Primary Type: primary description of the incident according to the Illinoise Uniform Crime Reporting code [https://data.cityofchicago.org/d/c7ck-438e.]

- Description: subcategory of the primary description

- Location Description: description of the location where the incident occurred

- Arrest: Whether an arrest was made

- Domestic: Whether the incident was domestic-related according to Illinois Domestic Violence Act

- Beat: Indicates the beat where the incident occurred. A beat is the smallest police geographic area – each beat has a dedicated police beat car. Three to five beats make up a police sector, and three sectors make up a police district. The Chicago Police Department has 22 police districts. See the beats at [https://data.cityofchicago.org/d/aerh-rz74.]

- District: Indicates the police district where the incident occurred. See the districts at [https://data.cityofchicago.org/d/fthy-xz3r.]

- Ward: The ward (City Council district) where the incident occurred. See the wards at [https://data.cityofchicago.org/d/sp34-6z76.]

- Community Area: Indicates the community area where the incident occurred. Chicago has 77 community areas. See the community areas at [https://data.cityofchicago.org/d/cauq-8yn6.]

- Latitude: The latitude of the location where the incident occurred.

- Longitude: The longitude of the location where the incident occurred.

- Year: Year the incident occurred.


```{r , cache=TRUE}
Crime = read.csv('Crime_2010_to_2017.csv')
```

```{r, cache=TRUE}
dim(Crime)
names(Crime)
head(Crime)
summary(Crime)
```


Code to clean data: convert 'true', 'false' values to 0 and 1, and date to date time format.

```
library(dplyr)
library(lubridate)
Crime = Crime %>% 
  mutate()
  filter(Year >= 2010) %>% 
  mutate(Arrest = as.numeric(Arrest == 'true'), Domestic = as.numeric(Domestic == 'true'), Date = as.POSIXct(strptime(Date, "%m/%d/%Y %H:%M:%S %p")))
head(Crime)
write.csv(Crime, file="Crime_2010_to_2017.csv")
```

# Demographics Data

### Description

Source of Data: We used the get_tract_demographics function from the acs package (American Community Survey) to get demographic data on Census Tracts from the US Census Bureau. The data spans the years 2010-2017.

Variables that we will use:

- region: the region id within a state

- total_population: the total population in the region

- percent_white: the percentage of people in the region that are white

- percent_black: the percentage of people in the region that are black

- percent_asian: the percentage of people in the region that are asian

- percent_hispanic: the percentage of people in the region that are hispanic

- per_capita_income: the capita per income of the region

- median_rent: the median rent of the region

- median_age: the median age of the region


```{r, cache=TRUE}
library(choroplethr)
library(choroplethrMaps)
library(acs)
api.key.install('23548fe7e771e8b0aeb4dd6e1ac95f7df1146a00')
illinois <- get_tract_demographics("illinois")
write.csv(illinois, 'Illinois_demographic.csv')
dim(illinois)
names(illinois)
head(illinois)
summary(illinois)
```


# Police Stops Data

```{r}
Police <- read.csv('Police_stops_1-2016_2-2017_full.csv')
head(Police)
dim(Police)

#Clean data to keep only a few variables
kept_columns <- c('CONTACT_CARD_ID', 'CONTACT_DATE', 'CONTACT_HOUR', 'SUBMITTING_BEAT_CD','SEX_CODE_CD','RACE_CODE_CD','ZIP_CD','BEAT','ENFORCEMENT_ACTION_TAKEN_I','COCAINE_I','HEROIN_I','CANNABIS_AMOUNT', 'OTHER_CON_SUB', 'SEARCH_COCAINE_I', 'SEARCH_HEROIN_I', 'SEARCH_CANNABIS_I', 'SEARCH_OTHER_CON_SUB_I', 'PARA_I', 'S_PARA_I')
drug_i <- c('COCAINE_I','HEROIN_I','CANNABIS_AMOUNT', 'OTHER_CON_SUB', 'SEARCH_COCAINE_I', 'SEARCH_HEROIN_I', 'SEARCH_CANNABIS_I', 'SEARCH_OTHER_CON_SUB_I', 'PARA_I', 'S_PARA_I', 'ENFORCEMENT_ACTION_TAKEN_I')
temp <- Police %>% select(one_of(kept_columns)) %>%
  filter(!grepl("REDACTED", RACE_CODE_CD), !(ZIP_CD == "" | ZIP_CD == " " | ZIP_CD == "null" | ZIP_CD == "NULL")) %>%
  na.omit() %>%
  mutate_at(drug_i, function(x){as.numeric(x == "Y")}) %>%
  mutate(DRUG_I = as.numeric(rowSums(.[10:19]) > 0))
temp.s <- temp[sample(1:nrow(temp), 10000,
                             replace=FALSE),]
write.csv(temp, file="Police_stops_2016_2017.csv")
write.csv(temp.s, file="Police_stops_2016_2017_sample.csv")
```