---
title: "Project Checkpoint 5"
author: "Jina Park, Phuc Nguyen"
date: "11/21/2017"
output: html_document
---

### Progress Made

### Group Roles

### Research Question 1

Predicting hotspots of specific crimes and how they changed over time.

Analyzing Hotspots of Crime Using a Bayesian Spatiotemporal Modeling Approach


**Models:**

The probability of an individual being a victim of a violent crime ($p_{ij}$)

### Research Question 2

Examining how crime rates have changed over time and differences in these changes across neighborhoods.

https://www.ml.cmu.edu/research/dap-papers/dap_flaxman.pdf

#### Clean data

```{r}

#Load data
Crime = read.csv('Crime_sample_2010_to_2017.csv')

Crime = Crime %>%
  mutate(Date = as.POSIXct(strptime(Date, "%Y-%m-%d %H:%M:%S"))) %>%
  filter(!is.na(Latitude))

#Project observations to ZIP codes boundaries

#load shapefile
zip_shp <- readOGR(dsn="Boundaries - ZIP Codes", layer="geo_export_e1262361-5c82-45ee-8427-ef228e06dc4a")
temp <- zip_shp
zip_df <- fortify(temp, region = "zip")

#needs to reassign CRS for shapefile
new_CRS <- CRS("+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
zip_shp <- sp::spTransform(zip_shp, new_CRS)

#summarize crime count per year
years <- unique(Crime$Year)
Hotspot_Crime <- data.frame()

for(i in 1:length(years)){
  year <- years[i]
  Current_Year <- Crime %>% filter(Year == year)
  
  locations <- with(Current_Year, as.data.frame(cbind(Longitude, Latitude)))
  coordinates(locations) <- ~Longitude+Latitude
  proj4string(locations) <- CRS("+init=epsg:4326")
      
  #prepping the data: project points to polygons
  proj4string(zip_shp) <- new_CRS
  proj4string(locations) <- new_CRS
  by_zip <- over(locations, zip_shp)
  
  by_zip <- by_zip %>%
    group_by(zip) %>%
    dplyr::summarise(total=n()) %>%
    filter(!is.na(zip)) %>%
    mutate(id = as.character(zip))
  
  total_map <- left_join(zip_df, by_zip)
  total_map <- total_map %>% mutate(Year = year)
  
  Hotspot_Crime <- rbind(Hotspot_Crime, total_map)
}


#Add demographic information
library(choroplethrZip)
data("df_zip_demographics")
Demographics <- df_zip_demographics %>% mutate(zip = region)
Hotspot_Crime <- inner_join(Hotspot_Crime, Demographics, by="zip") 
Model_Data <- Hotspot_Crime %>%
  select(-c(lat, long, order, piece, group, hole)) %>% distinct
```

```{r}
library(tidyr)
Year <- Model_Data %>%
  select(Year, zip, total) %>%
  spread( key=Year, value=total)
rownames(Year) <- Year$zip 
Year <- Year %>% select(-zip)
Year[is.na(Year)] <- 0

Population <- Model_Data %>%
  select(total_population, zip, Year) %>%
  spread(key=Year, value=total_population) 
rownames(Population) <- Population$zip
Population <- Population %>% select(-zip)
Population[is.na(Population)] <- 0

time <- sort(unique(Model_Data$Year))
```

####Simulation

```{r}
library(rjags)


#specify the model
crime_model <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i,j])
        logit(p[i,j]) = alpha + u[i] + s[i] + beta*time[j] + delta[i]*time[j]
      }
    }
    
    #Priors
    alpha ~ dunif(-1000, 1000)
    beta ~ dnorm(0, 1/1000)
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      s[i] ~ dnorm(0, prec.s)
      delta[i] ~ dnorm(0, prec.delta)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.s ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags <- jags.model(textConnection(crime_model), data=list(y=Year, n=Population, time=time),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim <- coda.samples(crime_jags, variable.names=c("p", "u", "s", "delta", "alpha", "beta"), n.iter=40000)


#store the samples in a data frame:    
crime_data <- data.frame(step=1:40000, crime_sim[[1]])
```

```{r}
summary(crime_sim)
```



