---
title: "Models Update"
author: "Jina Park, Phuc Nguyen"
date: "12/5/2017"
output: html_document
---

### Progress Made


### Group Roles


-----------------------------------------------------------

### Research Question 1

- Predicting hotspots of specific crimes and how they changed over time and space.

- Are certain types of crimes more prevalent in certain areas than others? 

- How have these hotspots changed over time? If they have changed over time, how has it changed relative to changes in income, demographics, age, etc?


### Research Question 2

- Examining police traffic stops and subsequent arrests to determine whether certain ethnic minorities are stopped/arrested more than others.

- Are certain ethnic groups more likely to be arrested/stopped in certain areas than others? For instance, are blacks/hispanics more likely to be stopped/arrested in predominantly white neighborhoods?

- We would like to examine the relationship between traffic stops/arrests and the actual crime rates of different ethnic groups. For instance, are blacks/hispanics getting stopped/arrested at a disproportionately greater rate than they are committing crimes? 

#### Load Libraries
```{r message=FALSE, warning=FALSE}
library(tidyr)
library(rjags)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(MacBayes)
library(choroplethrZip)
library(rgdal)
```

#### Map cleanup
```{r}
cleanup <- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = 'white', colour = 'white'),
        axis.line = element_line(colour = "white"),
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())
```

#### Clean data
```{r message=FALSE, warning=FALSE}
#Load data
Crime = read.csv('Crime_sample_2010_to_2017.csv')

Crime = Crime %>%
  mutate(Date = as.POSIXct(strptime(Date, "%Y-%m-%d %H:%M:%S"))) %>%
  filter(!is.na(Latitude))
```

#### Project observations to ZIP codes boundaries
```{r message=FALSE, warning=FALSE}
#load shapefile
zip_shp <- readOGR(dsn="Boundaries - ZIP Codes", layer="geo_export_e1262361-5c82-45ee-8427-ef228e06dc4a")
temp <- zip_shp
zip_df <- fortify(temp, region = "zip")

#needs to reassign CRS for shapefile
new_CRS <- CRS("+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
zip_shp <- sp::spTransform(zip_shp, new_CRS)

#summarize crime count per year
years <- unique(Crime$Year)
Hotspot_Crime <- data.frame()

for(i in 1:length(years)){
  year <- years[i]
  Current_Year <- Crime %>% filter(Year == year)
  
  locations <- with(Current_Year, as.data.frame(cbind(Longitude, Latitude)))
  coordinates(locations) <- ~Longitude+Latitude
  proj4string(locations) <- CRS("+init=epsg:4326")
      
  #prepping the data: project points to polygons
  proj4string(zip_shp) <- new_CRS
  proj4string(locations) <- new_CRS
  by_zip <- over(locations, zip_shp)
  
  by_zip <- by_zip %>%
    group_by(zip) %>%
    dplyr::summarise(total=n()) %>%
    filter(!is.na(zip)) %>%
    mutate(id = as.character(zip))
  
  total_map <- left_join(zip_df, by_zip)
  total_map <- total_map %>% mutate(Year = year)
  
  Hotspot_Crime <- rbind(Hotspot_Crime, total_map)
}
```

#### Add demographic information to Model Data
```{r message=FALSE, warning=FALSE}
data("df_zip_demographics")
Demographics <- df_zip_demographics %>% mutate(zip = region)
Hotspot_Crime <- inner_join(Hotspot_Crime, Demographics, by="zip") 
Model_Data <- Hotspot_Crime %>%
  select(-c(lat, long, order, piece, group, hole)) %>% distinct
```

#### Get populations by zipcode and year
```{r}
Crime_count <- Model_Data %>%
  select(Year, zip, total) %>%
  spread(key=Year, value=total)
rownames(Crime_count) <- Crime_count$zip 
Crime_count <- Crime_count %>% select(-zip)
Crime_count[is.na(Crime_count)] <- 0
head(Crime_count)
```

```{r}
Population <- Model_Data %>%
  select(total_population, zip, Year) %>%
  spread(key=Year, value=total_population) %>%
  select('2010', zip)
rownames(Population) <- Population$zip
Population <- Population %>% select(-zip)
Population[is.na(Population)] <- 0
head(Population)

Population.vector <- Population$`2010`
time <- sort(unique(Model_Data$Year))
percent_nonwhite <- 100 - Model_Data$percent_white
```

---------------------------------------------------------------

#### Simulation
# Research Question 1

**Plot of $log(y_{ij}/n_{i})$ vs year**
```{r}
Crime_count_log_calc <- log(Crime_count/Population.vector)

Crime_count_log <- Crime_count_log_calc %>%
  rownames_to_column() %>%
  gather(key=year, value=crime_count_log, -rowname) %>%
  mutate(zip=rowname) %>%
  select(zip, year, crime_count_log) %>%
  mutate(year=as.numeric(year))
head(Crime_count_log)

ggplot(Crime_count_log, aes(x=year, y=crime_count_log, colour=zip)) + geom_point() + geom_line() 
```

**Plot of $$log(y_{ij}/n_{i})$$ vs per capita income**
```{r}
Model_Data_1 <- Model_Data %>%
  filter(Year == 2010) %>%
  mutate(percent_nonwhite = 100 - percent_white) %>%
  select(zip, per_capita_income, percent_nonwhite) %>%
  right_join(Crime_count_log, by="zip") 

head(Model_Data_1)

ggplot(Model_Data_1, aes(x=log(per_capita_income), y=crime_count_log, colour=percent_nonwhite)) + geom_point() + labs(x="Log of Per Capita Income", y="Log of Crime Rate", title="Relationship between Per Capita Income and Log of Crime Rate")
```

**Plot of $log(y_{ij}/n_{i})$ vs nonwhite percentage**
```{r}
ggplot(Model_Data_1, aes(x=log(percent_nonwhite), y=crime_count_log)) + geom_point() + labs(x="Log of Nonwhite Percentage", y="Log of Crime Rate", title="Relationship between Nonwhite Percentage and Log of Crime Rate")
```

## Model 0
#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \alpha + \beta * time_{j}$$

#### Priors
$$\alpha \sim N(0, 1000^2)$$
$$\beta \sim N(0, 1000^2)$$

- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i]: the population in zipcode i

- $\beta$: time trend

- time[j]: the time in years

```{r}
#Data
Crime_count <- Model_Data %>%
  group_by(Year) %>%
  dplyr::summarise(total.stop = sum(total))
head(Crime_count)
```

```{r, cache=TRUE}
#Specify the model
crime_model_0 <- "model{
    #Data
    for(j in 1:8){
      y[j] ~ dbin(p[j], n)
      logit(p[j]) = beta*time[j]
    }
    
    #Priors
    beta ~ dnorm(0, 1/100^2)

}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_0 <- jags.model(textConnection(crime_model_0), data=list(y=Crime_count$total.stop, n=sum(Population.vector), time=Crime_count$Year),inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))
#crime_jags_0 <- jags.model(textConnection(crime_model_0), data=list(y=Crime_count$total.stop, n=sum(Population.vector), #time=Crime_count$Year),inits=list(alpha=-8, beta=-0.04))


#simulate a sample from the posterior 
crime_sim_0 <- coda.samples(crime_jags_0, variable.names=c("p", "beta"), n.iter=50000)


#store the samples in a data frame:    
crime_data_0 <- data.frame(crime_sim_0[[1]])
```

### Running Mean Plot
```{r}
running_mean_plot(crime_data_0$beta)
```

### Parameter Inferences
```{r}
ggplot(crime_data_0, aes(x=beta)) + geom_histogram(aes(y=..density..), colour="white")
mean(crime_data_0$beta)
```

### Confidence Intervals of Crime Model 0 parameters
```{r}
crime_model_0_CI <- matrix(nrow=9, ncol=2)
for (i in 1:9) {
      crime_model_0_CI[i,] <- quantile(crime_data_0[[i]], c(0.025, 0.975))
}
rownames(crime_model_0_CI) <- names(crime_data_0)
colnames(crime_model_0_CI) <- c("2.5%", "97.5%")
as.table(crime_model_0_CI)
```

## Model 1
### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

### Specify the Model
We are modeling the count of crime incidents by year and zipcodes using a binomial distribution, in which the parameters are the probability of being a victim and the total population in location i and year j.

We modeled the probability of being a victim using a logit function of the grand crime rate of Chicago and the inclusion of several random effects. It is also a function of time and its interaction with location.

#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \delta_{i} * time_{j}$$

#### Priors
$$\delta_{i} \sim N(0, prec.delta)$$

#### Hyperpriors
$$prec.delta \sim Gamma(0.5, 0.0005)$$

```{r, cache=TRUE}
#Data
Crime_count <- Model_Data %>%
  select(Year, zip, total) %>%
  spread(key=Year, value=total)
rownames(Crime_count) <- Crime_count$zip 
Crime_count <- Crime_count %>% select(-zip)
Crime_count[is.na(Crime_count)] <- 0
head(Crime_count)
```

```{r, cache=TRUE}
#Specify the model
crime_model_1 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i])
        logit(p[i,j]) = delta[i]*time[j]
      }
    }
    
    #Priors
    for(i in 1:58){
      delta[i] ~ dnorm(0, prec.delta)
    }

    #Hyperpriors
    prec.delta ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_1 <- jags.model(textConnection(crime_model_1), data=list(y=Crime_count, n=Population.vector, time=time),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


 #simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_1 <- coda.samples(crime_jags_1, variable.names=c("p", "delta"), n.iter=60000)


#store the samples in a data frame:    
crime_data_1 <- data.frame(crime_sim_1[[1]])
```

### Parameter Interpretation

```{r}
ggplot(crime_data_1, aes(x=alpha)) + geom_histogram(aes(y=..density..), colour="white")
sum(crime_data_1$alpha > 0)/40000
```

It appears that beta, the time trend, is slightly negative. This suggests that the amount of crime in Chicago might be decreasing over time.

```{r}
ggplot(crime_data_1, aes(x=beta)) + geom_histogram(aes(y=..density..), colour="white")
```

### Confidence Intervals of Crime Model 0 parameters
```{r}
crime_model_1_delta_CI <- matrix(nrow=length(names(crime_data_1)), ncol=2)
for (i in 1:length(names(crime_data_1))) {
      crime_model_1_delta_CI[i,] <- quantile(crime_data_1[[i]], c(0.025, 0.975))
}
rownames(crime_model_1_delta_CI) <- names(crime_data_1)
colnames(crime_model_1_delta_CI) <- c("2.5%", "97.5%")
as.table(crime_model_1_delta_CI)
```

The 95% confidence interval for $\beta$ is (-0.004096, 0.0002897). Since $\beta$ is not considered signifant in the 5% level, we can proceed with Model 2 instead, in which the overall time trend is removed from the model.

------------------------------------------------------------------------------

## Model 2 

**Overall time trend ($\beta$) is removed from the model**

### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- $\alpha$: the mean crime rate

- u[i]: the unstructured spatial random effect for each zipcode in Chicago.

- s[i]: the structured spatial random effect for each zipcode in Chicago.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

### Specify the Model
This model resembles Model 1 but we are no longer taking into account the time trend.

#### Data 
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \alpha + u_{i} + \delta_{i} * time_{j}$$

#### Priors
$$\alpha \sim N(0, 1000^2)$$
$$\beta \sim N(0, 1000^2)$$
$$u_{i} \sim N(0, prec.u)$$
$$\delta_{i} \sim N(0, prec.delta)$$

#### Hyperpriors
$$prec.u \sim Gamma(0.5, 0.0005)$$
$$prec.delta \sim Gamma(0.5, 0.0005)$$

```{r, cache=TRUE, message=FALSE, warning=FALSE}
#Specify the model
crime_model_2 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i])
        logit(p[i,j]) = alpha + u[i] + delta[i]*time[j]
      }
    }
    
    #Priors
    alpha ~ dnorm(0, 1/1000^2)
    beta ~ dnorm(0, 1/1000^2)
    
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      delta[i] ~ dnorm(0, prec.delta)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_2 <- jags.model(textConnection(crime_model_2), data=list(y=Crime_count, n=Population.vector, time=time),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_2 <- coda.samples(crime_jags_2, variable.names=c("p", "u", "delta", "alpha"), n.iter=40000)


#store the samples in a data frame:    
crime_data_2 <- data.frame(step=1:40000, crime_sim_2[[1]])
```

## Crime Model 2 Visualizations
```{r}
#Alter the data frame for visualizations
zip_data <- data.frame(zipcode=c(1:58), zip=rownames(Crime_count))

crime_viz_2 <- crime_data_2 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "delta")) %>%
  filter(grepl("delta", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
```

### Map of Crime Differential Growth by Zipcode (Using Model 2)
```{r}
zip <- readOGR(dsn="Boundaries - ZIP Codes", layer="geo_export_e1262361-5c82-45ee-8427-ef228e06dc4a")
zip_df <- fortify(zip, region = "zip")
crime_model_2_delta <- left_join(zip_df, crime_viz_2, by="id")
head(crime_model_2_delta)

model_2_map <- ggplot() 
model_2_map <- model_2_map +  geom_polygon(data = crime_model_2_delta, aes(x=long, y=lat, group=group, fill=delta), color = "black", size=0.2) 
model_2_map <- model_2_map + coord_map() 
model_2_map <- model_2_map + labs(x="", y="", title=paste("Crime Differential Growth by Zipcode"), fill="total")
model_2_map <- model_2_map + cleanup
print(model_2_map)
```

------------------------------------------------------------------------------

# Model 3 -- Use demographics information 

### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- $\alpha$: the mean crime rate

- u[i]: the unstructured spatial random effect for each zipcode in Chicago.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

- $\sigma$[i]: the income trend

- income[i]: income for each zipcode in Chicago

### Specify the Model
This model resembles Model 2 but now our prediction of the crime rate accounts for the different income levels in various Chicago zipcodes.

#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \alpha + u_{i} + \delta_{i} * time_{j} + \sigma_{i} * income_{i}$$

#### Priors
$$\alpha \sim N(0, 1000^2)$$
$$\beta \sim N(0, 1000^2)$$
$$u_{i} \sim N(0, prec.u)$$
$$\delta_{i} \sim N(0, prec.delta)$$
$$\sigma_{i} \sim N(0, prec.sigma)$$

#### Hyperprior
$$prec.u \sim Gamma(0.5, 0.0005)$$
$$prec.delta \sim Gamma(0.5, 0.0005)$$
$$prec.sigma \sim Gamma(0.5, 0.0005)$$

```{r, cache=TRUE, message=FALSE, warning=FALSE}
#Specify the model
crime_model_3 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i])
        logit(p[i,j]) = alpha + u[i] + delta[i] * time[j] + sigma[i] * income[i]
      }
    }
    
    #Priors
    alpha ~ dnorm(0, 1/1000^2)
    beta ~ dnorm(0, 1/1000^2)
    
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      delta[i] ~ dnorm(0, prec.delta)
      sigma[i] ~ dnorm(0, prec.sigma)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    prec.sigma ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_3 <- jags.model(textConnection(crime_model_3), data=list(y=Crime_count, n=Population.vector, time=time, income=Model_Data$per_capita_income),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_3 <- coda.samples(crime_jags_3, variable.names=c("p", "u", "delta", "alpha", "sigma"), n.iter=40000)


#store the samples in a data frame:    
crime_data_3 <- data.frame(step=1:40000, crime_sim_3[[1]])
```

## Crime Model 3 Visualizations
```{r}
crime_viz_3_delta <- crime_data_3 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "delta")) %>%
  filter(grepl("delta", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_3_delta <- left_join(zip_df, crime_viz_3_delta, by="id")
head(crime_model_3_delta)
```

```{r}
crime_viz_3_sigma <- crime_data_3 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "sigma")) %>%
  filter(grepl("sigma", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_3_sigma <- left_join(zip_df, crime_viz_3_sigma, by="id")
head(crime_model_3_sigma)
```

### Map of Crime Differential Growth by Zipcode (Using Model 3)
```{r}
model_3_map <- ggplot() 
model_3_map <- model_3_map +  geom_polygon(data = crime_model_3_delta, aes(x=long, y=lat, group=group, fill=delta), color = "black", size=0.2) 
model_3_map <- model_3_map + coord_map() 
model_3_map <- model_3_map + labs(x="", y="", title=paste("Crime Differential Growth by Zipcode"), fill="total")
model_3_map <- model_3_map + cleanup
print(model_3_map)
```

### Income Differential Growth by Zipcode (Using Model 3)
```{r}
model_3_map <- ggplot() 
model_3_map <- model_3_map +  geom_polygon(data = crime_model_3_sigma, aes(x=long, y=lat, group=group, fill=sigma), color = "black", size=0.2) 
model_3_map <- model_3_map + coord_map() 
model_3_map <- model_3_map + labs(x="", y="", title=paste("Income Differential Growth by Zipcode"), fill="total")
model_3_map <- model_3_map + cleanup
print(model_3_map)
```

In locations where the income differential growth increases, the crime differential growth seems to decrease.

### Running mean plot of alpha

The running mean of $\alpha$ appears to stabilize after 40,000 iterations.

```{r}
running_mean_plot(crime_data_3$alpha)
```

---------------------------------------------------------------

## Model 4 -- Use demographics information with race

### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- $\alpha$: the mean crime rate

- u[i]: the unstructured spatial random effect for each zipcode in Chicago.

- s[i]: the structured spatial random effect for each zipcode in Chicago.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

- $\sigma$[i]: the income trend

- income[i]: income for each zipcode in Chicago

- e[i]: the weight in the relationship between crime rate and percentage of nonewhite people in Chicago

- percent_nonwhite[i]: the percentage of nonwhite people in zipcode i.

### Specify the Model
This model resembles Model 3 but now our prediction of the crime rate accounts for both income levels and the percentage of nonwhites in the area.

#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \alpha + u_{i} + \delta_{i} * time_{j} + \sigma_{i} * income_{i} + \rho_{i} * percentNonwhite_{i}$$

#### Priors
$$\alpha \sim N(0, 1000^2)$$
$$\beta \sim N(0, 1000^2)$$
$$u_{i} \sim N(0, prec.u)$$
$$\delta_{i} \sim N(0, prec.delta)$$
$$\sigma_{i} \sim N(0, prec.sigma)$$
$$\rho_{i} \sim N(0, 0.01)$$

#### Hyperprior
$$prec.u \sim Gamma(0.5, 0.0005)$$
$$prec.delta \sim Gamma(0.5, 0.0005)$$
$$prec.sigma \sim Gamma(0.5, 0.0005)$$

```{r, cache=TRUE, message=FALSE, warning=FALSE}
#Specify the model
crime_model_4 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i])
        logit(p[i,j]) = alpha + u[i] + delta[i] * time[j] + sigma[i]*income[i] + rho[i]*percent_nonwhite[i]
      }
    }
    
    #Priors
    alpha ~ dnorm(0, 1/1000^2)
    beta ~ dnorm(0, 1/1000^2)

    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      delta[i] ~ dnorm(0, prec.delta)
      sigma[i] ~ dnorm(0, prec.sigma)
      rho[i] ~ dnorm(0, 0.01)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    prec.sigma ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_4 <- jags.model(textConnection(crime_model_4), data=list(y=Crime_count, n=Population.vector, time=time, income=Model_Data$per_capita_income, percent_nonwhite=100 - Model_Data$percent_white),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_4 <- coda.samples(crime_jags_4, variable.names=c("p", "u", "delta", "alpha", "sigma", "rho"), n.iter=40000)

#store the samples in a data frame:    
crime_data_4 <- data.frame(step=1:40000, crime_sim_4[[1]])
```

### Crime Model 4 Visualizations
```{r}
crime_viz_4_delta <- crime_data_4 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "delta")) %>%
  filter(grepl("delta", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_4_delta <- left_join(zip_df, crime_viz_4_delta, by="id")
head(crime_model_4_delta)
```

```{r}
crime_viz_4_sigma <- crime_data_4 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "sigma")) %>%
  filter(grepl("sigma", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_4_sigma <- left_join(zip_df, crime_viz_4_sigma, by="id")
head(crime_model_4_sigma)
```

```{r}
crime_viz_4_rho <- crime_data_4 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "rho")) %>%
  filter(grepl("rho", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_4_rho <- left_join(zip_df, crime_viz_4_rho, by="id")
head(crime_model_4_rho)
```

### Map of Crime Differential Growth by Zipcode (Using Model 4)
```{r}
model_4_map <- ggplot() 
model_4_map <- model_4_map +  geom_polygon(data = crime_model_4_delta, aes(x=long, y=lat, group=group, fill=delta), color = "black", size=0.2) 
model_4_map <- model_4_map + coord_map() 
model_4_map <- model_4_map + labs(x="", y="", title=paste("Crime Differential Growth by Zipcode"), fill="total")
model_4_map <- model_4_map + cleanup 
print(model_4_map)
```


### Map of Income Differential Growth by Zipcode (Using Model 4)
```{r}
model_4_map_sigma <- ggplot() 
model_4_map_sigma <- model_4_map_sigma +  geom_polygon(data = crime_model_4_sigma, aes(x=long, y=lat, group=group, fill=sigma), color = "black", size=0.2) 
model_4_map_sigma <- model_4_map_sigma + coord_map() 
model_4_map_sigma <- model_4_map_sigma + labs(x="", y="", title=paste("Income Differential Growth by Zipcode"), fill="total")
model_4_map_sigma <- model_4_map_sigma + cleanup
print(model_4_map_sigma)
```

### Map of Percentage Nonwhite Differential Growth by Zipcode (Using Model 4)
```{r}
model_4_map_rho <- ggplot() 
model_4_map_rho <- model_4_map_rho +  geom_polygon(data = crime_model_4_rho, aes(x=long, y=lat, group=group, fill=rho), color = "black", size=0.2) 
model_4_map_rho <- model_4_map_rho + coord_map() 
model_4_map_rho <- model_4_map_rho + labs(x="", y="", title=paste("Percentage Nonwhite Differential Growth by Zipcode"), fill="total")
model_4_map_rho <- model_4_map_rho + cleanup
print(model_4_map_rho)
```


------------------------------------------------------------------------


# Police Traffic Stops model

### Cleaning the data
#### Data frame for the total population by race and zipcode
```{r}
#Number of zipcodes: 62; Number of races: 7
Police <- read.csv("Police_stops_2016_2017_sample.csv")

Police.r <- Police %>%
  filter(!(RACE_CODE_CD=="I" | RACE_CODE_CD=="P" | RACE_CODE_CD=="WHI") ) %>%
  mutate(RACE_CODE_CD = as.character(RACE_CODE_CD)) %>%
  mutate(RACE_CODE_CD=replace(RACE_CODE_CD, RACE_CODE_CD=="WWH", "HISPANIC"))

vars = c("percent_white","percent_black","percent_asian","percent_hispanic")
Population.race <- Model_Data %>%
  mutate_at(vars, function(x){trunc(Model_Data$total_population*x/100)}) %>%
  select(zip, percent_asian,percent_black,percent_hispanic, percent_white) %>%
  unique() %>%
  remove_rownames %>%
  column_to_rownames(var='zip')
head(Population.race)
```

#### Data frame for the number of arrests by zipcode and race
```{r}
valid_zips <- unique(rownames(Population.race))
Arrests <- Police.r %>% 
  filter(ZIP_CD %in% valid_zips) %>%
  group_by(ZIP_CD, RACE_CODE_CD) %>%
  dplyr::summarise(total.arrest = as.integer(sum(ENFORCEMENT_ACTION_TAKEN_I))) %>%
  spread(key=ZIP_CD, value=total.arrest) %>%
  as.data.frame() %>%
  remove_rownames %>%
  column_to_rownames(var='RACE_CODE_CD')
Arrests[is.na(Arrests)] <- as.integer(0)
head(Arrests)
```

#### Data frame for the number of police stops by zipcode and race
```{r}
Stops <- Police.r %>%
  filter(ZIP_CD %in% valid_zips) %>%
  group_by(ZIP_CD, RACE_CODE_CD) %>%
  dplyr::summarise(total.stop = n()) %>%
  spread(key=ZIP_CD, value=total.stop) %>%
  as.data.frame() %>%
  remove_rownames %>%
  column_to_rownames(var='RACE_CODE_CD')
Stops[is.na(Stops)] <- 0
head(Stops)
```

-----------------------------------------------------------------

### Preliminary visualizations

#### Stop Rate by Race
```{r}
total_pop <- Model_Data %>% filter(Year==2010) %>% select(total_population)
population_by_ethnic <- Model_Data %>%
  filter(Year==2010) %>%
  mutate_at(c("percent_asian", "percent_black", "percent_hispanic", "percent_white"), function(x){x* total_pop$total_population/100}) %>%
  mutate(zip = as.numeric(zip)) %>%
  select(percent_asian, percent_black, percent_hispanic, percent_white, zip)
population_by_ethnic
print(colSums(population_by_ethnic))
data_stops = data.frame(race=c("Asian Pacific Islander", "Black", "Hispanic", "White"), stop_rate=(rowSums(Stops)/colSums(population_by_ethnic %>% select(-zip))))
rownames(data_stops) <- c()
head(data_stops %>% arrange(desc(stop_rate)))
```

Blacks and Hispanics are stopped at higher rate than Whites and Asians given their population.

### Plot of Traffic Stops by Population
```{r}
l <- population_by_ethnic %>%
  mutate(BLK=percent_black, HISPANIC=percent_hispanic, API=percent_asian, WHT=percent_white) %>%
  select(BLK, HISPANIC, API, WHT, zip) %>%
  gather(key=RACE_CODE_CD, value=population, - zip)

data <- Police.r %>%
  filter(ZIP_CD %in% valid_zips) %>%
  group_by(ZIP_CD, RACE_CODE_CD) %>%
  dplyr::summarise(total.stop = n()) %>%
  inner_join(l, by=c("ZIP_CD" = 'zip', 'RACE_CODE_CD')) %>%
  mutate(race=RACE_CODE_CD)

ggplot(data, aes(x=population, y=total.stop, colour=RACE_CODE_CD)) + geom_point() + geom_smooth() +
  labs(x="Population Count", y="Number of Traffic Stops", title="Traffic Stops By Population", subtitle="Visualization of number of traffic stops as the population of each ethnic group increases for each zipcode") + 
  guides(colour=guide_legend("Race")) +
  theme(plot.title = element_text(hjust=0.5)) 
```

### Plot of Traffic Stops by Percentage of Population
```{r}
ggplot(data, aes(x=population, y=total.stop, colour=RACE_CODE_CD)) + geom_point() + geom_smooth() +
  labs(x="Percent of Population", y="Number of Traffic Stops", fill="Race", title="Traffic Stops By Percentage of Population") + 
  guides(colour=guide_legend("Race")) +
  theme(plot.title = element_text(hjust=0.5))
```

------------------------------------------------------------


## Police Model 0

### Variables Used

- Data is only from January 2016 - February 2017 (13/12)

- i: ethnic group

- j: zipcode locations in Chicago

- y[i, j]: observed number of arrests of ethnic group i in area j

- n[i, j]: the number of traffic stops in area j for ethnic group i

- $\mu$: grand mean of arrests

- $\alpha$[i]: the random effect for each ethnic group i


#### Data 
$$y_{ij} \sim Pois(\frac{13}{12} n_{ij}  \theta_{ij})$$
$$log(\theta_{ij}) = \mu + \alpha_{i} $$

#### Priors
$$\mu \sim N(0, 0.000001)$$
$$\alpha_{i} \sim N(0, 0.000001)$$


```{r}
log(0.28603)
log(0.28603+0.05911) - log(0.28603)
log(0.28603-0.05712) - log(0.28603)
log(0.28603-0.12381) - log(0.28603)
```

```{r, cache=TRUE, message=FALSE, warning=FALSE}
police_model_0 <- "model{
  #Data
  for (i in 1:4) {
    for(j in 1:58){
      y[i,j] ~ dpois((13/12) * n[i,j] * theta[i,j])
      log(theta[i,j]) <- mu + alpha[i]
    }
  }

  #Priors
  mu ~ dnorm(0, 1.0E-06)

  for (i in 1:4){
    alpha[i] ~ dnorm(0, 0.005)
  }
}"

#initial value set
inits = list(mu=-1.251659, alpha=c(0, 0.1878534, -0.2227678, -0.5671433))

police_jags_0 <- jags.model(textConnection(police_model_0), data=list(y=Arrests, n=Stops),
                    inits=inits)

#simulate a sample from the posterior 
#note that we specify both mu and tau variables
police_sim_0 <- coda.samples(police_jags_0, variable.names=c("mu", "alpha"), n.iter=400000)

#store the samples in a data frame:    
police_data_0 <- data.frame(step=1:400000, police_sim_0[[1]])
```

```{r}
running_mean_plot(police_data_0$mu)
running_mean_plot(police_data_0$alpha.2.)
```

------------------------------------------------------------------

## Police Model 1

```{r}
#estimate initial values
data <- Police.r %>% 
  filter(ZIP_CD %in% valid_zips) %>%
  group_by(ZIP_CD, RACE_CODE_CD) %>%
  dplyr::summarise(total.arrest = as.integer(sum(ENFORCEMENT_ACTION_TAKEN_I)), total.stop = n())
head(data)
```

```{r}
freq_model <- lm(total.arrest ~ RACE_CODE_CD * total.stop, data = data)
summary(freq_model)
```

### Variables Used

- Data is only from January 2016 - February 2017 (13/12)

- i: ethnic group

- j: zipcode locations in Chicago

- y[i, j]: observed number of arrests of ethnic group i in area j

- n[i, j]: the number of traffic stops in area j for ethnic group i

- $\mu$: grand mean of arrests

- $\alpha$[i]: the random effect for each ethnic group i

- $\beta$[j]: the random effect for each zipcode location j

- $\epsilon$[i, j]: the random error for each ethnic group i and location j

#### Data 
$$y_{ij} \sim Pois(\frac{13}{12} n_{ij}  \theta_{ij})$$
$$log(\theta_{ij}) = \mu + \alpha_{i} + \beta_{j} + \epsilon_{ij}$$

#### Priors
$$\beta_{0} \sim N(0, 0.000001)$$
$$\alpha_{i} \sim N(0, 0.000001)$$
$$\beta_{j} \sim N(0, b^2)$$
$$\epsilon_{ij} \sim N(0, e^2)$$

#### Hyperpriors
$$b \sim Gamma(0.5, 0.0005)$$
$$e \sim Gamma(0.5, 0.0005)$$

```{r, cache=TRUE, message=FALSE, warning=FALSE}
police_model_1 <- "model{
  #Data
  for (i in 1:4) {
    for(j in 1:58){
      y[i,j] ~ dpois((13/12)*n[i,j]*theta[i,j])
      log(theta[i,j]) <- mu + alpha[i] + beta[j] + epsilon[i,j]
    }
  }

  #Priors

  mu ~ dnorm(0,1.0E-06)

  for (i in 1:4){
    alpha[i] ~ dnorm(0, 0.005)
  }

  for (j in 1:58){
    beta[j] ~ dnorm(0, 1/b^2)
  }

  for (i in 1:4){
    for (j in 1:58){
      epsilon[i,j] ~ dnorm(0, 1/e^2)
    }
  }

  b ~ dgamma(0.5, 0.0005)
  e ~ dgamma(0.5, 0.0005)
  
}"

#initial value set
inits = list(mu=-1.251659, alpha=c(0, 0.1878534, -0.2227678, -0.5671433), tau_beta=170, tau_epsilon=170, beta=rep(0.03, 58), epsilon=matrix(rep(0.01, 58*4), ncol=58))

police_jags_1 <- jags.model(textConnection(police_model_1), data=list(y=Arrests, n=Stops),
                    inits=inits)

#simulate a sample from the posterior 
#note that we specify both mu and tau variables
police_sim_1 <- coda.samples(police_jags_1, variable.names=c("mu", "alpha", "beta", "b", "e", "epsilon"), n.iter=500000)

#store the samples in a data frame:    
police_data_1 <- data.frame(step=1:500000, police_sim_1[[1]])
```

### Running mean plot of beta.1., mu, and alpha.2.
```{r}
running_mean_plot(police_data_1$beta.1.)
running_mean_plot(police_data_1$mu)
running_mean_plot(police_data_1$alpha.2.)
```


```{r}
police_viz <- police_data_1 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "beta")) %>%
  filter(grepl("beta", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  mutate(mu = mean(police_data_1$mu)) %>%
  mutate(rate = exp(mu + beta)) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
police_model_beta <- left_join(zip_df, police_viz, by="id")
head(police_model_beta)
```

### Map of rate at which traffic stops result in arrests by zipcode (Using Model 3)
```{r}
model_1_map_beta <- ggplot() 
model_1_map_beta <- model_1_map_beta +  geom_polygon(data = police_model_beta, aes(x=long, y=lat, group=group, fill=rate), color = "black", size=0.2) 
model_1_map_beta <- model_1_map_beta + coord_map() 
model_1_map_beta <- model_1_map_beta + labs(x="", y="", title=paste("Rate at Which Traffic Stops Result in Arrests by Zipcode"), fill="total")
model_1_map_beta <- model_1_map_beta + cleanup
print(model_1_map_beta)
```

### Rate of traffic stops by ethnic group
```{r}
police_viz1 <- police_data_1 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "alpha")) %>%
  filter(grepl("alpha", zipcode)) %>%
  mutate(mu = mean(police_data_1$mu)) %>%
  mutate(rate = exp(mu + alpha))
head(police_viz1)
```

