---
title: "Models Update"
author: "Jina Park, Phuc Nguyen"
date: "12/5/2017"
output: html_document
---

### Research Question 1

- Predicting hotspots of specific crimes and how they changed over time and space.

- Are certain types of crimes more prevalent in certain areas than others? 

- How have these hotspots changed over time? If they have changed over time, how has it changed relative to changes in income, demographics, age, etc?


### Research Question 2

- Examining police traffic stops and subsequent arrests to determine whether certain ethnic minorities are stopped/arrested more than others.

- Are certain ethnic groups more likely to be arrested/stopped in certain areas than others? For instance, are blacks/hispanics more likely to be stopped/arrested in predominantly white neighborhoods?

- We would like to examine the relationship between traffic stops/arrests and the actual crime rates of different ethnic groups. For instance, are blacks/hispanics getting stopped/arrested at a disproportionately greater rate than they are committing crimes? 

#### Load Libraries
```{r message=FALSE, warning=FALSE}
library(tidyr)
library(rjags)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(MacBayes)
library(choroplethrZip)
library(rgdal)
```

#### Map cleanup
```{r}
cleanup <- theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = 'white', colour = 'white'),
        axis.line = element_line(colour = "white"),
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())
```

#### Clean data
```{r}
#Load data
Crime = read.csv('Crime_sample_2010_to_2017.csv')

Crime = Crime %>%
  mutate(Date = as.POSIXct(strptime(Date, "%Y-%m-%d %H:%M:%S"))) %>%
  filter(!is.na(Latitude))
```

#### Project observations to ZIP codes boundaries
```{r message=FALSE, warning=FALSE}
#load shapefile
zip_shp <- readOGR(dsn="Boundaries - ZIP Codes", layer="geo_export_e1262361-5c82-45ee-8427-ef228e06dc4a")
temp <- zip_shp
zip_df <- fortify(temp, region = "zip")

#needs to reassign CRS for shapefile
new_CRS <- CRS("+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
zip_shp <- sp::spTransform(zip_shp, new_CRS)

#summarize crime count per year
years <- unique(Crime$Year)
Hotspot_Crime <- data.frame()

for(i in 1:length(years)){
  year <- years[i]
  Current_Year <- Crime %>% filter(Year == year)
  
  locations <- with(Current_Year, as.data.frame(cbind(Longitude, Latitude)))
  coordinates(locations) <- ~Longitude+Latitude
  proj4string(locations) <- CRS("+init=epsg:4326")
      
  #prepping the data: project points to polygons
  proj4string(zip_shp) <- new_CRS
  proj4string(locations) <- new_CRS
  by_zip <- over(locations, zip_shp)
  
  by_zip <- by_zip %>%
    group_by(zip) %>%
    dplyr::summarise(total=n()) %>%
    filter(!is.na(zip)) %>%
    mutate(id = as.character(zip))
  
  total_map <- left_join(zip_df, by_zip)
  total_map <- total_map %>% mutate(Year = year)
  
  Hotspot_Crime <- rbind(Hotspot_Crime, total_map)
}
```

#### Add demographic information to Model Data
```{r}
data("df_zip_demographics")
Demographics <- df_zip_demographics %>% mutate(zip = region)
Hotspot_Crime <- inner_join(Hotspot_Crime, Demographics, by="zip") 
Model_Data <- Hotspot_Crime %>%
  select(-c(lat, long, order, piece, group, hole)) %>% distinct
```


#### Matrix data: crime counts, populations by zipcode and year

#### Nested data: crime counts, population, standardized year 
```{r}
Model_Data <- transform(Model_Data, id=match(zip, unique(zip)))
Model_Data$time = Model_Data$Year-2009
Model_Data$racegr <- ifelse(Model_Data$percent_white >80, 1, 
                            ifelse(Model_Data$percent_white <=80 & Model_Data$percent_white > 60,2,
                            ifelse(Model_Data$percent_white <=60 & Model_Data$percent_white > 40,3,
                            ifelse(Model_Data$percent_white <=40 & Model_Data$percent_white > 20,4, 5)))) 
Model_Data$income = scale(Model_Data$per_capita_income)[,1]

write.csv(Model_Data, "Crime_models_data.csv")
```

### Load Data:
```{r}
Model_Data = read.csv("Crime_models_data.csv")
```


---------------------------------------------------------------

#### Simulation
# Research Question 1

**Plot of $log(y_{ij}/n_{i})$ vs year**
```{r, cache=TRUE}
Crime_count_log <- Model_Data %>%
  mutate(log_rate = log(total/total_population)) %>%
  select(zip, Year, log_rate)
head(Crime_count_log)

ggplot(Crime_count_log, aes(x=Year, y=log_rate, colour=zip)) + geom_point() + geom_line()  + labs(title="Crime Rate Over Time", y="Log of Crime Rate", x="Year", color="Zipcodes")
```

**Plot of $log(y_{ij}/n_{i})$ vs per capita income**
```{r}
Model_Data_1 <- Model_Data %>%
  filter(Year == 2010) %>%
  mutate(percent_nonwhite = 100 - percent_white) %>%
  select(zip, per_capita_income, percent_nonwhite) %>%
  right_join(Crime_count_log, by="zip") 
head(Model_Data_1)

ggplot(Model_Data_1, aes(x=log(per_capita_income), y=log_rate, colour=percent_nonwhite)) + geom_point() + labs(x="Log of Per Capita Income", y="Log of Crime Rate", title="Relationship between Per Capita Income and Crime Rate", color="Nonwhite Percentage") 
```

**Plot of $log(y_{ij}/n_{i})$ vs nonwhite percentage**
```{r}
ggplot(Model_Data_1, aes(x=log(percent_nonwhite), y=log_rate, colour=per_capita_income)) + geom_point() + labs(x="Log of Nonwhite Percentage", y="Log of Crime Rate", title="Relationship between Nonwhite Percentage and Log of Crime Rate")
```

## Model 0
#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \beta * time_{j}$$

#### Priors
$$\beta \sim N(0, 1000^2)$$

- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i]: the population in zipcode i

- $\beta$: time trend

- time[j]: the time in years


```{r, cache=TRUE}
#Specify the model
crime_model_0 <- "model{
    #Data
    for(j in 1:ns){
      y[j] ~ dbin(p[j], n[j])
      logit(p[j]) = alpha + beta*time[j]
    }
    
    #Priors
    beta ~ dnorm(0, 1/1000^2)
    alpha ~ dnorm(0, 1/1000^2)

}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_0 <- jags.model(textConnection(crime_model_0), data=list(y=Model_Data$total, n=Model_Data$total_population, time=Model_Data$time, ns=nrow(Model_Data)),inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
crime_sim_0 <- coda.samples(crime_jags_0, variable.names=c("beta", "alpha"), n.iter=20000)


#store the samples in a data frame:    
crime_data_0 <- data.frame(crime_sim_0[[1]])
```


### Check for convergence
```{r}
#check for convergece
#from: https://rpubs.com/corey_sparks/32494
init.rng1<-list(".RNG.seed" = 1234, ".RNG.name" = "base::Mersenne-Twister")
init.rng2<-list(".RNG.seed" = 5678, ".RNG.name" = "base::Mersenne-Twister")
mod<-jags.model(file=textConnection(crime_model_0),list(y=Model_Data$total, n=Model_Data$total_population, time=Model_Data$time, ns=nrow(Model_Data)),inits =list(init.rng1, init.rng2) , n.chains=2)
update(mod, 10000)
samps<-coda.samples(mod, variable.names=c("beta","alpha"), n.iter=5000, n.thin=5)
gelman.diag(samps)

#running mean plot
running_mean_plot(crime_data_0$beta)
running_mean_plot(crime_data_0$alpha)
```

### Parameter Inferences
#### Posterior Distribution Plots
```{r}
ggplot(crime_data_0, aes(x=beta)) + geom_histogram(aes(y=..density..), colour="white") + labs(title="Posterior Distribution of the Effect of Time", x="Time Effect", y="Density")

ggplot(crime_data_0, aes(x=alpha)) + geom_histogram(aes(y=..density..), colour="white") + labs(title="Posterior Distribution of Intercept", x="Time Effect", y="Density")
```

#### Confidence Intervals and Posterior Means of Crime Model 0 Parameters
```{r}
crime_model_0_CI <- matrix(nrow=2, ncol=2)

for (i in 1:2) {
      crime_model_0_CI[i,] <- quantile(crime_data_0[[i]], c(0.025, 0.975))
}

crime_model_0_mean <- matrix(nrow=2, ncol=1)
for (i in 1:2) {
      crime_model_0_mean[i, ] <- mean(crime_data_0[[i]])
}

crime_model_0_table <- cbind(crime_model_0_CI, crime_model_0_mean)
rownames(crime_model_0_table) <- names(crime_data_0)
colnames(crime_model_0_table) <- c("2.5%", "97.5%", "Mean")
# as.table(crime_model_0_table)
```

## Model 1
### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

### Specify the Model
We are modeling the count of crime incidents by year and zipcodes using a binomial distribution, in which the parameters are the probability of being a victim and the total population in location i and year j.

We modeled the probability of being a victim using a logit function of the grand crime rate of Chicago and the inclusion of several random effects. It is also a function of time and its interaction with location.

#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \beta_{i} + \delta_{i} * time_{j}$$

#### Priors
$$\beta_{i} \sim N(0, \tau_{0})$$
$$\delta_{i} \sim N(0, \tau_{1})$$

#### Hyperpriors
$$\tau_0 \sim Gamma(0.5, 0.0005)$$

$$\tau_1 \sim Gamma(0.5, 0.0005)$$

```{r, cache=TRUE}
#Data
data = list(y=Model_Data$total, n=Model_Data$total_population, time=Model_Data$time, zipid=Model_Data$id, ns=length(Model_Data$total), nzip=length(unique(Model_Data$id)))
```

```{r, cache=TRUE}
#Specify the model
crime_model_1 <- "model{
    #Data
    for(i in 1:ns) {
      y[i] ~ dbin(p[i], n[i])
      logit(p[i]) = beta[zipid[i]] + delta[zipid[i]]*time[i]
    }
    
    #Priors
    for(j in 1:nzip){
      delta[j] ~ dnorm(0, tau_1)
      beta[j] ~ dnorm(0, tau_0)
    }

    #Hyperpriors
    tau_0 ~ dgamma(0.5, 0.0005)
    tau_1 ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior
#set the random number seed
init.rng1<-list(".RNG.seed" = 1998, ".RNG.name" = "base::Mersenne-Twister")
crime_jags_1 <- jags.model(textConnection(crime_model_1), data=data,
                    inits=init.rng1)


#simulate a sample from the posterior 
crime_sim_1 <- coda.samples(crime_jags_1, variable.names=c("delta", "beta", "tau_0", "tau_1"), n.iter=30000)


#store the samples in a data frame:    
crime_data_1 <- data.frame( crime_sim_1[[1]])
```

### Check for convergence
```{r}
#check for convergece
#from: https://rpubs.com/corey_sparks/32494
init.rng1<-list(".RNG.seed" = 1234, ".RNG.name" = "base::Mersenne-Twister")
init.rng2<-list(".RNG.seed" = 5678, ".RNG.name" = "base::Mersenne-Twister")
mod<-jags.model(file=textConnection(crime_model_1), data=data,inits =list(init.rng1, init.rng2) , n.chains=2)
update(mod, 10000)
samps<-coda.samples(mod, variable.names=c("delta", "alpha", "tau_delta", "tau_alpha"), n.iter=5000, n.thin=5)
gelman.diag(samps)


running_mean_plot(crime_data_1$delta.1.)
running_mean_plot(crime_data_1$beta.1.)
running_mean_plot(crime_data_1$tau_0)
running_mean_plot(crime_data_1$tau_1)
```

### Parameter Interpretation
#### Posterior Distribution
```{r}
ggplot(crime_data_1, aes(x=delta.1.)) + geom_histogram(aes(y=..density..), colour="white") + labs(title="Posterior Distribution of the Crime Differential Growth in 2010", x="Crime Differential Growth", y="Density")
```

### Confidence Intervals and Posterior Means of Crime Model 1 Parameters
```{r}
n = length(unique(Model_Data$zip))

#posterior mean and quantile of delta only

crime_model_1_CI <- matrix(nrow=n, ncol=2)
for (i in 1:n) {
      crime_model_1_CI[i, ] <- quantile(crime_data_1[[i+n]], c(0.025, 0.975))
}

crime_model_1_mean <- matrix(nrow=n, ncol=1)
for (i in 1:n) {
      crime_model_1_mean[i, ] <- mean(crime_data_1[[i+n]])
}

crime_model_1_table <- cbind(crime_model_1_CI, crime_model_1_mean)
rownames(crime_model_1_table) <- names(crime_data_1)[59:116]
colnames(crime_model_1_table) <- c("2.5%", "97.5%", "Mean")
# as.table(crime_model_1_table)
```

The 95% confidence interval for $\beta$ is (-0.004096, 0.0002897). Since $\beta$ is not considered signifant in the 5% level, we can proceed with Model 2 instead, in which the overall time trend is removed from the model.

## Crime Model 1 Visualizations

### Map of Crime Differential Growth by Zipcode (Using Model 1)
```{r, cache=TRUE}
zip <- readOGR(dsn="Boundaries - ZIP Codes", layer="geo_export_e1262361-5c82-45ee-8427-ef228e06dc4a")
zip_df <- fortify(zip, region = "zip")
crime_model_1_delta <- left_join(zip_df, data.frame(id=as.character(unique(Model_Data$zip)), delta=crime_model_1_mean), by='id')
head(crime_model_1_delta)

model_1_map <- ggplot() 
model_1_map <- model_1_map +  geom_polygon(data = crime_model_1_delta, aes(x=long, y=lat, group=group, fill=delta), color = "black", size=0.2) 
model_1_map <- model_1_map + coord_map() 
model_1_map <- model_1_map + labs(x="", y="", title=paste("Crime Differential Growth by Zipcode"), fill="total")
model_1_map <- model_1_map + cleanup
print(model_1_map)
```

### Variability of the Crime Differential Growth
```{r}
sqrt(1/mean(crime_data_1$tau_1))
```

```{r, cache=TRUE}
ggplot(crime_model_1_delta, aes(x=delta)) + geom_histogram(colour="white") + labs(x="Crime Differential Growth", y="Count", title="Histogram of Crime Differential Growth")
```

------------------------------------------------------------------------------

# Model 3 -- Use demographics information 

### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

- $\sigma$[i]: the income trend

- income[i]: income for each zipcode in Chicago

### Specify the Model
This model resembles Model 2 but now our prediction of the crime rate accounts for the different income levels in various Chicago zipcodes.

#### Data

$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) = \beta_{i} + \sigma_{j} * income_{i}$$

#### Priors
$$\beta_{i} \sim N(0, \tau_{0})$$
$$\sigma_{j} \sim N(0, \tau_{1})$$

#### Hyperprior

$$\tau_{0} \sim Gamma(0.5, 0.0005)$$
$$\tau_{1} \sim Gamma(0.5, 0.0005)$$

```{r}
Model3_Data <- Model_Data %>% filter(Year==2010)
data = list(y=Model3_Data$total, n=Model3_Data$total_population, zipid=Model3_Data$id, racegr=Model3_Data$racegr, income=Model3_Data$income, ns=nrow(Model3_Data), nzip=length(unique(Model3_Data$zip)))
```

```{r, cache=TRUE, message=FALSE, warning=FALSE}
#Specify the model
crime_model_3 <- "model{
    #Data
    for(i in 1:ns) {
      y[i] ~ dbin(p[i], n[i])
      logit(p[i]) = beta[racegr[i]] + sigma[racegr[i]]*income[i]
    }
    
    #Priors
    for(j in 1:5){
      sigma[j] ~ dnorm(0, tau_1)
      beta[j] ~ dnorm(0, tau_0)
    }

    #Hyperpriors
    tau_0 ~ dgamma(0.5, 0.0005)
    tau_1 ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_3 <- jags.model(textConnection(crime_model_3), data=data,
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_3 <- coda.samples(crime_jags_3, variable.names=c("sigma", "tau_0", "beta","tau_1"), n.iter=30000, n.thin=5)


#store the samples in a data frame:    
crime_data_3 <- data.frame(crime_sim_3[[1]])
```

```{r}
running_mean_plot(crime_data_3$beta.1.)
running_mean_plot(crime_data_3$sigma.1.)
running_mean_plot(crime_data_3$tau_0)
running_mean_plot(crime_data_3$tau_1)
```

## Crime Model 3 Visualizations

```{r}
posterior_means <- crime_data_3 %>% colMeans()
posterior_means
crime_model_3_sigma <- Model_Data
crime_model_3_sigma$beta <- recode(crime_model_3_sigma$racegr, `1`=posterior_means[1],
                                   `2`=posterior_means[2],
                                   `3`=posterior_means[3],
                                   `4`=posterior_means[4],
                                   `5`=posterior_means[5])
crime_model_3_sigma$sigma <- recode(crime_model_3_sigma$racegr, `1`=posterior_means[6],
                                   `2`=posterior_means[7],
                                   `3`=posterior_means[8],
                                   `4`=posterior_means[9],
                                   `5`=posterior_means[10])
crime_model_3_sigma <- crime_model_3_sigma %>%
  mutate(id = as.character(zip)) %>%
  right_join(zip_df, by="id")
```


### Income Differential Growth by Zipcode (Using Model 3)
```{r, cache=TRUE}
model_3_map <- ggplot() 
model_3_map <- model_3_map +  geom_polygon(data = crime_model_3_sigma, aes(x=long, y=lat, group=group, fill=sigma), color = "black", size=0.2) 
model_3_map <- model_3_map + coord_map() 
model_3_map <- model_3_map + labs(x="", y="", title=paste("Crime Differential Growth Based on Income and Demographics"), fill="total")
model_3_map <- model_3_map + cleanup
print(model_3_map)

posterior_means
#Five groups, 20% increase in percent white. Darkest blue is the areas with 20% or lower of population white.
#See Posterior means
```

In locations where the income differential growth increases, the crime differential growth seems to decrease.

### Running mean plot of alpha

The running mean of $\alpha$ appears to stabilize after 40,000 iterations.

---------------------------------------------------------------

## Model 4 -- Use demographics information with race

### Notation
- i: zipcodes in Chicago (58 total in our data)

- j: years (from 2010 to 2017)

- y[i, j]: observed count of violent crimes in Chicago by year and zipcode

- p[i, j]: probability of being a victim in location i and time j

- n[i, j]: the population in zipcode i in time j.

- time[j]: the time in years

- $\delta$[i]: the spatial temporal interaction term

- $\sigma$[i]: the income trend

- income[i]: income for each zipcode in Chicago

- $\rho$[i]: the weight in the relationship between crime rate and percentage of nonewhite people in Chicago

- percent_nonwhite[i]: the percentage of nonwhite people in zipcode i.

### Specify the Model
This model resembles Model 3 but now our prediction of the crime rate accounts for both income levels and the percentage of nonwhites in the area.

#### Data
$$y_{ij} \sim Bin(p_{ij}, n_{i})$$
$$logit(p_{ij}) =  \delta_{i} * time_{j} + \sigma_{i} * income_{i} + \rho_{i} * percentNonwhite_{i}$$

#### Priors
$$\delta_{i} \sim N(0, \tau_{1})$$
$$\sigma_{i} \sim N(0, tau_{2})$$
$$\rho_{i} \sim N(0, 0.01)$$

#### Hyperprior
$$\tau_{1} \sim Gamma(0.5, 0.0005)$$
$$\tau_{2} \sim Gamma(0.5, 0.0005)$$
$$\tau_{3} \sim Gamma(0.5, 0.0005)$$


{r, cache=TRUE, message=FALSE, warning=FALSE}
#Specify the model
crime_model_4 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i])
        logit(p[i,j]) = delta[i] * time[j] + sigma[i]*income[i] + rho[i]*percent_nonwhite[i] + gamma[i]*income[i]*percent_nonwhite[i]
      }
    }
    
    #Priors
    for(i in 1:58){
      delta[i] ~ dnorm(-0.005, tau1)
      sigma[i] ~ dnorm(0, tau2)
      rho[i] ~ dnorm(0, 0.01)
      gamma[i] ~ dnorm(0, tau3)
    }

    #Hyperpriors
    tau1 ~ dgamma(0.005, 0.005)
    tau2 ~ dgamma(0.005, 0.005)
    tau3 ~ dgamma(0.005, 0.005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_4 <- jags.model(textConnection(crime_model_4), data=list(y=Crime_count, n=Population.vector, time=time, income=Model_Data$per_capita_income, percent_nonwhite=percent_nonwhite),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_4 <- coda.samples(crime_jags_4, variable.names=c("p", "delta", "sigma", "rho", "tau1", "tau2", "tau3"), n.iter=60000)

#store the samples in a data frame:    
crime_data_4 <- data.frame(crime_sim_4[[1]])
```

### Running Mean Plots
{r}
running_mean_plot(crime_data_4$delta.4.)
running_mean_plot(crime_data_4$sigma.4.)
running_mean_plot(crime_data_4$rho.4.)
```

### Crime Model 4 Visualizations
{r}
crime_viz_4_delta <- crime_data_4 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "delta")) %>%
  filter(grepl("delta", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_4_delta <- left_join(zip_df, crime_viz_4_delta, by="id")
head(crime_model_4_delta)
```

{r}
crime_viz_4_sigma <- crime_data_4 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "sigma")) %>%
  filter(grepl("sigma", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_4_sigma <- left_join(zip_df, crime_viz_4_sigma, by="id")
head(crime_model_4_sigma)
```

{r}
crime_viz_4_rho <- crime_data_4 %>%
  colMeans() %>%
  as.data.frame() %>%
  rownames_to_column(var="zipcode") %>%
  setNames(., c("zipcode", "rho")) %>%
  filter(grepl("rho", zipcode)) %>%
  mutate(zipcode = as.integer(gsub("\\D", "", zipcode))) %>%
  left_join(zip_data, by="zipcode") %>%
  mutate(id=as.character(zip))
crime_model_4_rho <- left_join(zip_df, crime_viz_4_rho, by="id")
head(crime_model_4_rho)
```

### Map of Crime Differential Growth by Zipcode (Using Model 4)
{r, cache=TRUE}
model_4_map <- ggplot() 
model_4_map <- model_4_map +  geom_polygon(data = crime_model_4_delta, aes(x=long, y=lat, group=group, fill=delta), color = "black", size=0.2) 
model_4_map <- model_4_map + coord_map() 
model_4_map <- model_4_map + labs(x="", y="", title=paste("Crime Differential Growth by Zipcode"), fill="total")
model_4_map <- model_4_map + cleanup 
print(model_4_map)
```


### Map of Income Differential Growth by Zipcode (Using Model 4)
{r, cache=TRUE}
model_4_map_sigma <- ggplot() 
model_4_map_sigma <- model_4_map_sigma +  geom_polygon(data = crime_model_4_sigma, aes(x=long, y=lat, group=group, fill=sigma), color = "black", size=0.2) 
model_4_map_sigma <- model_4_map_sigma + coord_map() 
model_4_map_sigma <- model_4_map_sigma + labs(x="", y="", title=paste("Income Differential Growth by Zipcode"), fill="total")
model_4_map_sigma <- model_4_map_sigma + cleanup
print(model_4_map_sigma)
```

### Map of Percentage Nonwhite Differential Growth by Zipcode (Using Model 4)
{r, cache=TRUE}
model_4_map_rho <- ggplot() 
model_4_map_rho <- model_4_map_rho +  geom_polygon(data = crime_model_4_rho, aes(x=long, y=lat, group=group, fill=rho), color = "black", size=0.2) 
model_4_map_rho <- model_4_map_rho + coord_map() 
model_4_map_rho <- model_4_map_rho + labs(x="", y="", title=paste("Percentage Nonwhite Differential Growth by Zipcode"), fill="total")
model_4_map_rho <- model_4_map_rho + cleanup
print(model_4_map_rho)
```

