---
title: "Project Checkpoint 5"
author: "Jina Park, Phuc Nguyen"
date: "11/21/2017"
output: html_document
---

### Progress Made

- Examined other research papers with similar topics to derive models for our research questions.

- Produced 2 models for each of our research questions.

### Group Roles

-----------------------------------------------------------

### Research Question 1

Predicting hotspots of specific crimes and how they changed over time.

Analyzing Hotspots of Crime Using a Bayesian Spatiotemporal Modeling Approach


**Models:**

The probability of an individual being a victim of a violent crime ($p_{ij}$)

### Research Question 2

Examining how crime rates have changed over time and differences in these changes across neighborhoods.

https://www.ml.cmu.edu/research/dap-papers/dap_flaxman.pdf

#### Clean data
```{r}
#Load data
Crime = read.csv('Crime_sample_2010_to_2017.csv')

Crime = Crime %>%
  mutate(Date = as.POSIXct(strptime(Date, "%Y-%m-%d %H:%M:%S"))) %>%
  filter(!is.na(Latitude))
```


#### Project observations to ZIP codes boundaries
```{r}
#load shapefile
zip_shp <- readOGR(dsn="Boundaries - ZIP Codes", layer="geo_export_e1262361-5c82-45ee-8427-ef228e06dc4a")
temp <- zip_shp
zip_df <- fortify(temp, region = "zip")

#needs to reassign CRS for shapefile
new_CRS <- CRS("+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
zip_shp <- sp::spTransform(zip_shp, new_CRS)

#summarize crime count per year
years <- unique(Crime$Year)
Hotspot_Crime <- data.frame()

for(i in 1:length(years)){
  year <- years[i]
  Current_Year <- Crime %>% filter(Year == year)
  
  locations <- with(Current_Year, as.data.frame(cbind(Longitude, Latitude)))
  coordinates(locations) <- ~Longitude+Latitude
  proj4string(locations) <- CRS("+init=epsg:4326")
      
  #prepping the data: project points to polygons
  proj4string(zip_shp) <- new_CRS
  proj4string(locations) <- new_CRS
  by_zip <- over(locations, zip_shp)
  
  by_zip <- by_zip %>%
    group_by(zip) %>%
    dplyr::summarise(total=n()) %>%
    filter(!is.na(zip)) %>%
    mutate(id = as.character(zip))
  
  total_map <- left_join(zip_df, by_zip)
  total_map <- total_map %>% mutate(Year = year)
  
  Hotspot_Crime <- rbind(Hotspot_Crime, total_map)
}
```


#### Add demographic information to Model Data
```{r}
library(choroplethrZip)
data("df_zip_demographics")
Demographics <- df_zip_demographics %>% mutate(zip = region)
Hotspot_Crime <- inner_join(Hotspot_Crime, Demographics, by="zip") 
Model_Data <- Hotspot_Crime %>%
  select(-c(lat, long, order, piece, group, hole)) %>% distinct
```

```{r}
library(tidyr)
Year <- Model_Data %>%
  select(Year, zip, total) %>%
  spread( key=Year, value=total)
rownames(Year) <- Year$zip 
Year <- Year %>% select(-zip)
Year[is.na(Year)] <- 0

Population <- Model_Data %>%
  select(total_population, zip, Year) %>%
  spread(key=Year, value=total_population) 
rownames(Population) <- Population$zip
Population <- Population %>% select(-zip)
Population[is.na(Population)] <- 0

time <- sort(unique(Model_Data$Year))

percent_nonwhite <- 100 - Model_Data$percent_white
```

---------------------------------------------------------------

#### Simulation
```{r}
library(rjags)
```

# Research Question 1

## Model 1
```{r}
#Specify the model
crime_model_1 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i,j])
        logit(p[i,j]) = alpha + u[i] + s[i] + beta*time[j] + delta[i]*time[j]
      }
    }
    
    #Priors
    alpha ~ dunif(-1000, 1000)
    beta ~ dnorm(0, 1/1000)
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      s[i] ~ dnorm(0, prec.s)
      delta[i] ~ dnorm(0, prec.delta)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.s ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_1 <- jags.model(textConnection(crime_model_1), data=list(y=Year, n=Population, time=time),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


 #simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_1 <- coda.samples(crime_jags_1, variable.names=c("p", "u", "s", "delta", "alpha", "beta"), n.iter=60000)


#store the samples in a data frame:    
crime_data_1 <- data.frame(step=1:60000, crime_sim_1[[1]])
```


## Model 2 - Overall time trend ($\beta$) is removed from the model
```{r}
#Specify the model
crime_model_2 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i,j])
        logit(p[i,j]) = alpha + u[i] + s[i]+ delta[i]*time[j]
      }
    }
    
    #Priors
    alpha ~ dunif(-1000, 1000)
    beta ~ dnorm(0, 1/1000)
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      s[i] ~ dnorm(0, prec.s)
      delta[i] ~ dnorm(0, prec.delta)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.s ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_2 <- jags.model(textConnection(crime_model_2), data=list(y=Year, n=Population, time=time),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_2 <- coda.samples(crime_jags_2, variable.names=c("p", "u", "s", "delta", "alpha"), n.iter=40000)


#store the samples in a data frame:    
crime_data_2 <- data.frame(step=1:40000, crime_sim_2[[1]])
```


# Model 3 -- Use demographics information 

```{r}
#Specify the model
crime_model_3 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i,j])
        logit(p[i,j]) = alpha + u[i] + s[i]+ delta[i]*time[j] + sigma[i]*income[i]
      }
    }
    
    #Priors
    alpha ~ dunif(-1000, 1000)
    beta ~ dnorm(0, 1/1000)
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      s[i] ~ dnorm(0, prec.s)
      delta[i] ~ dnorm(0, prec.delta)
      sigma[i] ~ dnorm(0, prec.sigma)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.s ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    prec.sigma ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_3 <- jags.model(textConnection(crime_model_3), data=list(y=Year, n=Population, time=time, income=Model_Data$per_capita_income),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_3 <- coda.samples(crime_jags_3, variable.names=c("p", "u", "s", "delta", "alpha", "sigma"), n.iter=40000)


#store the samples in a data frame:    
crime_data_3 <- data.frame(step=1:40000, crime_sim_3[[1]])
```

```{r}
ggplot(crime_data_3, aes(x=alpha)) + geom_histogram(aes(y=..density..))
```


```{r}
ggplot(crime_data_3, aes(x=step, y=alpha)) + geom_line()
```

```{r}
library(MacBayes)
running_mean_plot(crime_data_3$alpha)
```

# Model 4 -- Use demographics information with race

```{r}
#Specify the model
crime_model_4 <- "model{
    #Data
    for(i in 1:58) {
      for(j in 1:8){
        y[i,j] ~ dbin(p[i,j], n[i,j])
        logit(p[i,j]) = alpha + u[i] + s[i]+ delta[i]*time[j] + sigma[i]*income[i] + e[i]*percent_nonwhite[i]
      }
    }
    
    #Priors
    alpha ~ dunif(-1000, 1000)
    beta ~ dnorm(0, 1/1000)
    for(i in 1:58){
      u[i] ~ dnorm(0, prec.u)
      s[i] ~ dnorm(0, prec.s)
      delta[i] ~ dnorm(0, prec.delta)
      sigma[i] ~ dnorm(0, prec.sigma)
      e[i] ~ dnorm(0, 0.01)
    }

    #Hyperpriors
    prec.u ~ dgamma(0.5, 0.0005)
    prec.s ~ dgamma(0.5, 0.0005)
    prec.delta ~ dgamma(0.5, 0.0005)
    prec.sigma ~ dgamma(0.5, 0.0005)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model (nba_model) and data (y)
#set the random number seed
crime_jags_4 <- jags.model(textConnection(crime_model_4), data=list(y=Year, n=Population, time=time, income=Model_Data$per_capita_income, percent_nonwhite=percent_nonwhite),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
crime_sim_4 <- coda.samples(crime_jags_4, variable.names=c("p", "u", "s", "delta", "alpha", "sigma", "e"), n.iter=40000)


#store the samples in a data frame:    
crime_data_4 <- data.frame(step=1:40000, crime_sim_4[[1]])
```

#Police model


```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
#number of zipcodes: 62
#number races 7
Police <- read.csv("Police_stops_2016_2017_sample.csv")

Police.r <- Police %>%
  filter(!(RACE_CODE_CD=="I" | RACE_CODE_CD=="P" | RACE_CODE_CD=="WHI") ) %>%
  mutate(RACE_CODE_CD = as.character(RACE_CODE_CD)) %>%
  mutate(RACE_CODE_CD=replace(RACE_CODE_CD, RACE_CODE_CD=="WWH", "HISPANIC"))

valid_zips <- unique(rownames(Population.race))

Arrests <- Police.r %>% 
  filter(ZIP_CD %in% valid_zips) %>%
  group_by(ZIP_CD, RACE_CODE_CD) %>%
  dplyr::summarise(total.arrest = sum(ENFORCEMENT_ACTION_TAKEN_I)) %>%
  spread(key=ZIP_CD, value=total.arrest) %>%
  as.data.frame() %>%
  remove_rownames %>%
  column_to_rownames(var='RACE_CODE_CD')
Arrests[is.na(Arrests)] <- 0

Stops <- Police.r %>%
  filter(ZIP_CD %in% valid_zips) %>%
  group_by(ZIP_CD, RACE_CODE_CD) %>%
  dplyr::summarise(total.stop = n()) %>%
  spread(key=ZIP_CD, value=total.stop) %>%
  as.data.frame() %>%
  remove_rownames %>%
  column_to_rownames(var='RACE_CODE_CD')
Stops[is.na(Stops)] <- 0

vars = c("percent_white","percent_black","percent_asian","percent_hispanic")
Population.race <- Model_Data %>%
  mutate_at(vars, function(x){trunc(Model_Data$total_population*x/100)}) %>%
  select(zip, percent_asian,percent_black,percent_hispanic, percent_white) %>%
  unique() %>%
  remove_rownames %>%
  column_to_rownames(var='zip') 
```

## Model 1

```
#Specify the model
police_model_1 <- "model{
    #Data
    for(i in 1:4) {
      for(j in 1:58){
        y[i,j] ~ dpois((13/12)*n[i,j]*theta[i,j])
        log(theta[i,j]) = mu + alpha[i] + beta[j] + epsilon[i,j]
      }
    }
    
    #Priors
    mu ~ dunif(-500, 500)

    for (j in 1: 58){
      beta[j] ~ dnorm(0, 1/sigma_beta^2)
    }

    for (i in 1:4){
      alpha[i] ~ dunif(-50, 50)
    }

    for (i in 1:4){
      for (j in 1:62){
        epsilon[i,j] ~ dnorm(0, 1/sigma_epsilon^2)
      }
    }

    #Hyperpriors
    sigma_beta ~ dunif(0.00001, 0.01)
    sigma_epsilon ~ dunif(0.00001, 0.01)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model and data (y)
#set the random number seed
police_jags_1 <- jags.model(textConnection(police_model_1), data=list(y=Stops, n=Arrests),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
police_sim_1 <- coda.samples(police_jags_1, variable.names=c("mu", "alpha", "sigma_beta", "sigma_epsilon", "beta", "epsilon", "theta"), n.iter=10000)


#store the samples in a data frame:    
police_data_1 <- data.frame(step=1:10000, police_sim_1[[1]])
```

##Model 2

```
#Specify the model
police_model_2 <- "model{
    #Data
    for(i in 1:4) {
      for(j in 1:58){
        y[i,j] ~ dpois((13/12)*theta[i,j]*exp(alpha))
        n[i,j] ~ dpois(theta[i,j])
        log(theta[i,j]) = log(N[j,i]) + beta
      }
    }
    
    #Priors
    alpha ~ dnorm(0, 0.000001)
    beta ~ dnorm(0, 0.000001)
    
}"

#set up an algorithm to simulate the posterior by 
#combining the model and data (y)
#set the random number seed
police_jags_2 <- jags.model(textConnection(police_model_2), data=list(n=Stops, y=Arrests, N=Population.race),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
police_sim_2 <- coda.samples(police_jags_2, variable.names=c("alpha",  "beta",  "theta"), n.iter=10000)


#store the samples in a data frame:    
police_data_2 <- data.frame(step=1:10000, police_sim_2[[1]])
```

# Model 3

```{r, cache=TRUE}
police_model_3 <- "model{
  for (i in 1:4) {
    for(j in 1:58){
      Y[i,j] ~ dpois(lambda[i,j])
      log(lambda[i,j]) <- beta0 + (beta1 + beta2[i])*X[i,j]
    }
  }

  beta0 ~ dnorm(0,1.0E-06)
  beta1 ~ dnorm(0,1.0E-06)
  for (i in 1:4){
    beta2[i] ~ dnorm(0,1.0E-06)
  }
}"


police_jags_3 <- jags.model(textConnection(police_model_3), data=list(Y=Arrests, X=Stops),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
police_sim_3 <- coda.samples(police_jags_3, variable.names=c( "beta0", "beta1", "beta2"), n.iter=20000)


#store the samples in a data frame:    
police_data_3 <- data.frame(step=1:20000, police_sim_3[[1]])
```

```{r}
plot(police_sim_3)
```


# Model 4

```{r, cache=TRUE}
police_model_4 <- "model{
  for (i in 1:4) {
    for(j in 1:58){
      Y[i,j] ~ dpois(lambda[i,j])
      log(lambda[i,j]) <- beta0 + beta2[i] + beta3[j] + beta1*X[i,j]
    }
  }

  beta0 ~ dnorm(0,1.0E-06)
  beta1 ~ dnorm(0,1.0E-06)
  for (i in 1:4){
    beta2[i] ~ dnorm(0,1.0E-06)
  }
  for (j in 1: 58){
    beta3[j] ~ dnorm(0,1.0E-06)
  }

}"


police_jags_4 <- jags.model(textConnection(police_model_4), data=list(Y=Arrests, X=Stops),
                    inits=list(.RNG.name="base::Wichmann-Hill", .RNG.seed=1989))


#simulate a sample from the posterior 
#note that we specify both mu and tau variables
police_sim_4 <- coda.samples(police_jags_4, variable.names=c( "beta0", "beta1", "beta2", "beta3"), n.iter=20000)


#store the samples in a data frame:    
police_data_4 <- data.frame(step=1:20000, police_sim_4[[1]])
```


```{r}
running_mean_plot(police_data_4$beta0)
running_mean_plot(police_data_4$beta1)
running_mean_plot(police_data_4$beta2.2)
```